### 2017-11-12 Maths Notes/M2/Collisions/Collisions.tex
Updated notes to current date (12/11/17)

\begin{document}
\begin{center}
\underline{\huge Collisions}
\end{center}
\begin{obeylines}
\section{Impulse and momentum}
Impulse=$mv-mu$=$Ft$\\
Total momentum before collision=total momentum after\\
\section{Coefficient of restitution}
This tells us how well something bounces, it is given the symbol \textbf{e}.\\
If $e=1$ the ball returns to it's original height\\
If $e=0$ the ball doesn't bounce\\
$e=\dfrac{\text{Speed of seperation}}{\text{Speed of approach}}$

\end{obeylines}
\end{document}

>>>>

\begin{document}
\begin{center}
\underline{\huge Collisions}
\end{center}
\section{Impulse and momentum}
Impulse=$mv-mu$=$Ft$\\
Total momentum before collision=total momentum after
\section{Coefficient of restitution}
This tells us how well something bounces, it is given the symbol \textbf{e}.\\
If $e=1$ the ball returns to it's original height\\
If $e=0$ the ball doesn't bounce\\
\\
$e=\dfrac{\text{Speed of seperation}}{\text{Speed of approach}}$
\subsection{Alternate form of coefficient of restitution formula}
$mgh=\frac{1}{2}mv^2$\\
\\
$v=\sqrt{2gh}$\\
\\
$e=\dfrac{\sqrt{2gh_2}}{\sqrt{2gh_1}}$\\
\\
$e=\dfrac{\sqrt{h_2}}{\sqrt{h_1}}$\\
\\
$h_2$ - the height the ball bounces back to\\
\\
$h_1$ - the height the ball is dropped from
\subsection{Calculations involving coefficient of restitution}
When doing calculations involving the coefficient of restitution both the calculation for CoR and conservation of momentum will be needed.\\
\textbf{Conservation of momentum:}
$m_1u_1+m_2u_2=m_1v_1+m_2v_2$\\
\\
\textbf{Coefficient of restitution}\\
\\
$e=\dfrac{v_1}{u_1}$
\subsection{Successive Impacts}
In some cases calculations will involve the impacts of multiple balls successively, like in newton's cradle.
\subsubsection{Example}
\textit{Three perfectly elastic particles A, B and C with masses 3kg, 2kg and 1kg respectively lie at rest in a straight line on a smooth horizontal table in alphabetical order. A is projected towards B with speed $5ms^{-1}$ and after A has collided with B, B collides with C}\\
\\
$5\times3=3V_A+2V_B$\\
\\
$1=\dfrac{SoS}{SoA}$ therefore $SoS=SoA$ so $V_B-V_A=5$ so $V_B=5+V_A$\\
\\
$15=10+2V_A+3V_A$\\
\\
$15=2V_B+3$ so after 1st collision $V_B=6$\\
\\
$15=10+5V_A$ so $\mathbf{V_A=1}$\\
\\
\textbf{2nd Collision}\\
\\
$2\times6=2V_B+V_C$\\
\\
$V_C-V_B=6$\\
\\
$12=3V_B+6$\\
\\
$3V_B=6$\\
\\
$\mathbf{V_B=2}$\\
\\
$V_C=6+2=8$

\end{document}

### 2017-11-12 Maths Notes/S2/Hypothesis testing/Hypothesis testing.tex
Updated notes to current date (12/11/17)

The \textbf{boundary values} of the critical region are called the critical values.
\section{Method}
\begin{enumerate}
\item Establish the null and alternative hypothesis ($H_0$ and $H_1$)
\item Define distribution under $H_0$
\item Decide on the significance level
\item Collect data, state the test statistic, X=
\item Calculate the probability of obtaining the test statistic or a more extreme result (same direction as $H_1$)
\item Compare this to the sig level as a decimal
\begin{itemize}
\item If \textbf{greater} than the sig level, it is a \textbf{non significant} result, it is not in the critical region and we \textbf{do not} reject $H_0$
\item If \textbf{less} than sig level, it is a \textbf{significant result}, it is in the critical region and we \textbf{reject} $H_0$ 
\end{itemize}
\item Interpret the results in terms of the original claim
\end{enumerate}
\subsection{Example}
\textbf{Establish the null and alternative hypothesis}\\
$H_0:p=0.5$\\
$H_1:p>0.5$\\
\\
\textbf{Define the distribution under $\mathbf{H_O}$}\\
Under $H_0$ $X\sim B(15,0.5)$\\
\\
\textbf{Decide on the significance level}\\
$5\%$\\
\\
\textbf{Collect data, state the test statistic}\\
X=12\\
\\
\textbf{Calculate the probability of obtaining the test statistic or a more extreme result}\\
\begin{tabular}{r l}
$P(X\geqslant 12)$&$=1-P(X\leqslant 11)$\\
&$=1-0.9824$\\
&$=0.0176$\\
\end{tabular}
\\
\textbf{Compare this to the sig level as a decimal}
$0.0176<0.05$\\
\\
\textbf{Interpret the results in terms of the original claim}\\
There is evidence to reject $H_0$ in favour of $H_1$. The test is significant.
\subsection{Finding critical values}
We require a value c such that:\\
$P(X\geqslant c)<0.05$\\
$1-P(X\leqslant c-1)<0.05$\\
$P(X\leqslant c-1)>0.95$\\
\\
\textbf{Test against tables}\\
$P(X<11)=0.9824$\\
\\
$c-1=11$\\
$c=12$
\subsection{Two tailed tests}

>>>>

The \textbf{boundary values} of the critical region are called the critical values.
\section{Method}
\begin{enumerate}
\item Establish the null and alternative hypothesis ($H_0$ and $H_1$)
\item Define distribution under $H_0$
\item Decide on the significance level
\item Collect data, state the test statistic, X=
\item Calculate the probability of obtaining the test statistic or a more extreme result (same direction as $H_1$)
\item Compare this to the sig level as a decimal
\begin{itemize}
\item If \textbf{greater} than the sig level, it is a \textbf{non significant} result, it is not in the critical region and we \textbf{do not} reject $H_0$
\item If \textbf{less} than sig level, it is a \textbf{significant result}, it is in the critical region and we \textbf{reject} $H_0$ 
\end{itemize}
\item Interpret the results in terms of the original claim
\end{enumerate}
\subsection{Example}
\textbf{Establish the null and alternative hypothesis}\\
$H_0:p=0.5$\\
$H_1:p>0.5$\\
\\
\textbf{Define the distribution under $\mathbf{H_O}$}\\
Under $H_0$ $X\sim B(15,0.5)$\\
\\
\textbf{Decide on the significance level}\\
$5\%$\\
\\
\textbf{Collect data, state the test statistic}\\
X=12\\
\\
\textbf{Calculate the probability of obtaining the test statistic or a more extreme result}\\
\begin{tabular}{r l}
$P(X\geqslant 12)$&$=1-P(X\leqslant 11)$\\
&$=1-0.9824$\\
&$=0.0176$\\
\end{tabular}
\\
\textbf{Compare this to the sig level as a decimal}
$0.0176<0.05$\\
\\
\textbf{Interpret the results in terms of the original claim}\\
There is evidence to reject $H_0$ in favour of $H_1$. The test is significant.
\subsection{Finding critical values}
We require a value c such that:\\
$P(X\geqslant c)<0.05$\\
$1-P(X\leqslant c-1)<0.05$\\
$P(X\leqslant c-1)>0.95$\\
\\
\textbf{Test against tables}\\
$P(X<11)=0.9824$\\
\\
$c-1=11$\\
$c=12$
\subsection{Two tailed tests}
When doing a two tailed test the significance level must be split in two. For example if the significance level is 5\% then it must be split into 2.5\% for each tail.\\
\\
Two tailed test also concern equal or not equal, rather than inequalities, for example:\\
$H_0 : p=0.15$\\
$H_1 : p\neq0.15$

### 2017-11-12 Physics Notes/Capacitors/Capacitors.tex
Updated notes to current date (12/11/17)

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{circuitikz}
\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

@@

When charging the current flows round the top half of the circuit, when discharging the SPDT is changed to the other output and the current flows round the bottom half of the circuit.\\
\\
Capacitors can be used as simple timers in circuits.
\\
\\
The time it takes to charge/discharge depends on:
\begin{itemize}
\item The capacitance of the capacitor \textbf{(C)}
\item The resistance of the charging/discharging circuit \textbf{(R)}
\end{itemize}
\section{Energy stored in a capacitor}
\begin{tikzpicture}[scale=0.5]
\begin{axis}[xmin=0,ymin=-0.01,ylabel=V,xlabel=Q,axis lines=middle,label style={font=\LARGE},
ytick={2},yticklabels=v,
ticklabel style = {font=\LARGE},
xtick={1.5,2.5},xticklabels={q,q+$\Delta$q}
]
\addplot[mark=none]{x};
\addplot [mark=none] coordinates {(1.5,0) (1.5, 1.5)};
\addplot [mark=none] coordinates {(2.5,0) (2.5, 2.5)};
\addplot[mark=none,domain=0:2]{2};
\end{axis}
\filldraw[black, ultra thin,fill=gray!70] (2.5,0) -- (4.2,0) -- (4.2,3.5) -- (2.5,2.1) -- cycle;
\end{tikzpicture}\\
Area under graph=work done\\
V is the average P.D. as charge increases from q to q+$\Delta$q\\
$\Delta \text{w=v}\Delta \text{q}$\\
\\
$Q=\frac{1}{2}QV=E$  \textbf{(Work done=energy stored)}\\
\\
How to answer the exam question: Show that the energy stored by a capacitor is given by $E=\frac{1}{2}QV$
\begin{enumerate}
\item Sketch a graph of Q against V and describe what it shows
\item Describe that electrical energy is the product of charge and voltage
\item State QV is represented by the area under the line
\item The area under the line is a triangle with area =$\frac{1}{2}QV$
\item Therefore $E=\frac{1}{2}QV$
\end{enumerate}
$E=\frac{1}{2}QV$\\
$Q=CV$\\
Therefore:\\
$E=\frac{1}{2}CV^2$\\
$E=\frac{1}{2}\dfrac{Q^2}{C}$\\
\\
The battery supplies energy QV to the circuit but the capacitor only stores $\frac{1}{2}QV$, this means that 50\% of the energy provided by the battery is wasted due to resistance in the circuit.

\end{document}

>>>>

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{circuitikz}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage{relsize}
\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

@@

When charging the current flows round the top half of the circuit, when discharging the SPDT is changed to the other output and the current flows round the bottom half of the circuit.\\
\\
Capacitors can be used as simple timers in circuits.
\\
\\
The time it takes to charge/discharge depends on:
\begin{itemize}
\item The capacitance of the capacitor \textbf{(C)}
\item The resistance of the charging/discharging circuit \textbf{(R)}
\end{itemize}
\section{Energy stored in a capacitor}
\begin{tikzpicture}[scale=0.5]
\begin{axis}[xmin=0,ymin=-0.01,ylabel=V,xlabel=Q,axis lines=middle,label style={font=\LARGE},
ytick={2},yticklabels=v,
ticklabel style = {font=\LARGE},
xtick={1.5,2.5},xticklabels={q,q+$\Delta$q}
]
\addplot[mark=none]{x};
\addplot [mark=none] coordinates {(1.5,0) (1.5, 1.5)};
\addplot [mark=none] coordinates {(2.5,0) (2.5, 2.5)};
\addplot[mark=none,domain=0:2]{2};
\end{axis}
\filldraw[black, ultra thin,fill=gray!70] (2.5,0) -- (4.2,0) -- (4.2,3.5) -- (2.5,2.1) -- cycle;
\end{tikzpicture}\\
Area under graph=work done\\
V is the average P.D. as charge increases from q to q+$\Delta$q\\
$\Delta \text{w=v}\Delta \text{q}$\\
\\
$Q=\frac{1}{2}QV=E$  \textbf{(Work done=energy stored)}\\
\\
How to answer the exam question: Show that the energy stored by a capacitor is given by $E=\frac{1}{2}QV$
\begin{enumerate}
\item Sketch a graph of Q against V and describe what it shows
\item Describe that electrical energy is the product of charge and voltage
\item State QV is represented by the area under the line
\item The area under the line is a triangle with area =$\frac{1}{2}QV$
\item Therefore $E=\frac{1}{2}QV$
\end{enumerate}
$E=\frac{1}{2}QV$\\
$Q=CV$\\
Therefore:\\
$E=\frac{1}{2}CV^2$\\
$E=\frac{1}{2}\dfrac{Q^2}{C}$\\
\\
The battery supplies energy QV to the circuit but the capacitor only stores $\frac{1}{2}QV$, this means that 50\% of the energy provided by the battery is wasted due to resistance in the circuit.
\section{Charging and discharging a capacitor}
\begin{tikzpicture}[scale=0.5]
\begin{axis}[ticks=none,ylabel=Voltage\textbackslash V,xlabel=Time\textbackslash S,axis lines=middle,
label style={font=\large},xmin=0,xmax=2, 
 x label style={at={(axis description cs:0.5,0)},anchor=north,font=\LARGE},
 y label style={at={(axis description cs:0,.5)},rotate=90,anchor=south,font=\LARGE},
]
\addplot[mark=none,domain=0:2]{1/exp(x)};
\end{axis}
\end{tikzpicture}\\
The smaller the resistor connected to the capacitor, the steeper the initial gradient\\
\\
$\text{p.d.}\propto \text{rate of change of p.d.}$\\
\\
\textbf{Formula for discharge of a capacitor}\\
$\mathlarger{\mathlarger{Q=Q_0e^{-\frac{t}{RC}}}}$\\
\\
Where:\\
Q=Charge at time, t\\
$Q_0$=Initial charge\\
\\
$\mathlarger{\mathlarger{e^{-\frac{t}{RC}}}}$=Multiplication factor between 0 and 1\\
\\
This also works for:\\
\\
$\mathlarger{\mathlarger{V=V_0e^{-\frac{t}{RC}}}}$\\
\textbf{Formula for charging of a capacitor}\\
$\mathlarger{\mathlarger{Q=Q_0(1-e^{-\frac{t}{RC}})}}$\\
\\
\textbf{Rearranging for time}\\
\\
$\mathlarger{\mathlarger{V=V_0e^{-\frac{t}{RC}}}}$\\
\\
$\dfrac{V}{V_0}=e^{-\frac{t}{RC}}$\\
$\ln\Big(\dfrac{V}{V_0}\Big)=-\dfrac{t}{RC}$\\
$t=-RC\ln\Big(\dfrac{V}{V_0}\Big)$\\
$t=RC\ln\Big(\dfrac{V_0}{V}\Big)$\\
\\
\subsection{Proof of the units of the time constant being seconds}
$\tau=RC$\\\\
$\tau=\dfrac{V}{I}\times\dfrac{Q}{V}=\dfrac{Q}{I}$\\
\\
$\dfrac{Q}{I}=t$(unit seconds)
\end{document}

### 2017-11-12 Physics Notes/Further Mechanics and Thermal Physics/Thermal Physics/Thermal Physics.tex
Updated notes to current date (12/11/17)

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{tikz}
\usetikzlibrary{scopes}
\pgfplotsset{width=10cm,compat=1.9}

@@

\begin{document}
\begin{center}
\underline{\huge Circular Motion}
\end{center}
\section{Introduction to circular motion}
For an object to be going in a circle a force is needed to accelerate the object. This is because in a circle an object is constantly changing velocity, even if it isn't changing speed.\\
Centripetal force is not an extra force, it is just the name for a resultant force.\\
For an object to be deflected towards the centre, there needs to be a force acting towards the centre.
\subsection{Definitions}
\begin{tabular}{|c|c|c|c|}
\hline
Quantity&Symbol&Unit&Definition\\
\hline
Period&T&Seconds,s&The time taken for an object to complete one revolution\\
\hline
Frequency&f&Hertz,hz&The number of revolutions per second\\
\hline
Angular velocity&$\omega$&$\textrm{rads}^{-1}$&The rate of change in angle(the angle covered per second)\\
\hline
\end{tabular}
\subsection{Formula not on formula book}
$\omega=\dfrac{\Delta\theta}{\Delta t}$\\
$a=\dfrac{\Delta V}{\Delta t}$\\
Don't forget application of  $F=ma$ with the equations for force
\section{Circular motion for going over a hump}
\begin{tikzpicture}[
    force/.style={>=latex,draw=black,fill=black},
    m/.style={rectangle,draw=black,fill=white,minimum size=0.3cm,thin},
]
    \node[m] (m) {};
    {[force,->]
        \draw (m.north) -- ++(0,1) node[above] {$r$};
        \draw (m.south) -- ++(0,-1) node[below] {$mg$};
    }
\end{tikzpicture}
\\
For an object going over a hump the force of the weight is in the opposite direction to the reaction force.\\
The object will leave the ground if $mg>r$.
\subsection{Calculations}
$F_c=mg-R=\dfrac{mv^2}{r}$\\
\textbf{Calculating the speed at which the car will leave the ground (R=0)}\\
$mg=\dfrac{mv_0^2}{r} \qquad V_0=\sqrt{gr}$\\
Once the car has left the ground, use SUVAT equations rather than circular motion.
\section{Roundabouts}
On a roundabout the centripetal force is provided by friction.\\
$F_r=\dfrac{mv_0^2}{r}$\\
To avoid slipping, the centripetal force must be less than the maximum possible for the car and road surface.\\
\textbf{Example question}
\textit{The maximum possible force of friction between a car and the road is 800N, given a mass of 1000kg and a radius of 5m, what is the maximum speed the car can go without slipping}
\newpage
\section{Banked motion}
N is the reaction force\\
$n\sin\theta=\dfrac{mv^2}{r}$ - Force towards the centre of notion is equal to the centripetal force\\
$N\cos\theta=mg$ - The force vertically downwards is equal to the weight\\
$\tan\theta=\dfrac{v^2}{rg}$ - Combining the equations\\
$v^2=gr\tan\theta$ - Re-write
\section{Forces in a circle}
\begin{tikzpicture}[
    force/.style={>=latex,draw=black,fill=black},
    m/.style={rectangle,draw=black,fill=lightgray,minimum size=0.3cm,thin},
]
    \node[m] (m) at (0,-2) {};
    \node[m] (m2) at (0,2) {};
    \node[m] (m3) at (2,0) {};
    {[force,->]
        \draw (m.north) -- ++(0,0.7) node[above] {$R$};
        \draw (m.south) -- ++(0,-0.7) node[right] {$mg$};
    }
    {[force,->]
        \draw (m2.south) -- ++(0,-0.7) node[below] {$R+mg$};
    }
    {[force,->]
        \draw (m3.west) -- ++(-0.7,0) node[left] {$R$};
        \draw (m3.south) -- ++(0,-0.7) node[below right] {$mg$};
    }

\begin{pgfonlayer}{background layer}
    \draw[color=black, very thick](0,0) circle (2);
    \end{pgfonlayer}{background layer}
\end{tikzpicture}
\\
Top: Centripetal force = $R+W$\\
Side: Centripetal force = $R$\\
Bottom: Centripetal force = $R-W$\\
At the top of the loop r=0\\
$v=\sqrt{gr}$

>>>>

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{tikz}
\usetikzlibrary{scopes}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage{gensymb}

@@

\begin{document}
\begin{center}
\underline{\huge Thermal Physics}
\end{center}
\section{Differences between heat and temperature}
\begin{tabular}{|c|c|c|}
\hline
&Heat&Temperature\\
\hline
Definition&Thermal energy(transferred from hot to cooler places)&A comparative measure of how hot something is\\
\hline
Unit&Joule&Kelvin\\
\hline
Measured using&Joulemeter&Thermometer\\
\hline
\end{tabular}
\section{Graph of heating water}
\includegraphics[width=3in]{water_heat_curve.jpg}
\section{Specific heat capacity}
\textbf{Specific heat capacity} - The energy needed to raise the temperature of 1kg of a material by 1K\\
\\
$c=\dfrac{Q}{m\Delta\theta}$\\
\\
c=Specific heat capacity - $Jkg^{-1} \ \degree C$\\
m=Mass - $kg$\\
$\Delta\theta$ = Temperature change - $\degree C$\\
Q = Heat energy - $J$
\subsection{Latent heat}
\textbf{Specific latent heat of fusion, $\mathbf{L_f}$} \quad \textcolor{red}{$Q=mL_f$}\\
The energy needed to change 1kg of a solid to a liquid without a temperature change\\
\\
\textbf{Specific latent heat of vaporisation, $\mathbf{L_v}$} \quad \textcolor{red}{$Q=mL_v$}\\
The energy needed to change 1kg of a liquid to a vapour without a temperature change
